---
title: "Preprocessing and Comaprison of Sample Filtering -- CD66b-GR-Status Differential Expression"
author: "Elana Fertig & Joe Tandurella"
date: '2022-06-24'
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "~/OneDrive - Johns Hopkins/JHU_PDAC_Collaboration/Analysis_HTMLs/") })
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R libraries and versions

```{r}
library('readxl')
library('DESeq2')
library('ComplexHeatmap')
library('EnhancedVolcano')
library('pca3d')
library('sva')
library('DT')
library('fgsea')
library('GSVA')
library(ggplot2)
library(ggpubr)
library(tidyverse)
library(dplyr)
library(tidyr)
library(data.table)
set.seed(1234)
sessionInfo()
```

# Read in GEO data

## Expression counts data

```{r}
load('NewGeneCounts.Rda')
exprsDat <- txi.tx.genes.all$counts
```

## Sample Annotations

```{r}
sampAnnot <- as.data.frame(read_xls('j1568 sorted rna sample_attributes_survival_immuno_annot_LeiEdited.xls'))
sampAnnot$Treatment[sampAnnot$Treatment == 'GVAX_CY_Nivo'] <- 
  'GVAX_CY_NIVO'
rownames(sampAnnot) <- make.names(sampAnnot$Sample_id)
```

# Data normalization and exploratory data analysis

## Limit data to treatment groups of interest

```{r}
sampAnnot <- sampAnnot[sampAnnot$`Processed as` != 'control',]
sampAnnot <- sampAnnot[sampAnnot$Treatment != "GVAX_SBRT_Pembro",]

sampAnnot$batch <- ifelse(sapply(strsplit(sampAnnot$Sample_id, split="_"),length) == 4, 'A', 'B')

sampAnnot[sampAnnot$batch=='A','Sample_id'] <- 
  substring(sampAnnot$Sample_id,4)[sampAnnot$batch=='A']

exprsDat <- exprsDat[,rownames(sampAnnot)]

sampAnnot <- sampAnnot[colnames(exprsDat),]
```

## Evaluation of sample quality based upon distribution of reads

We evaluate sample quality from the distribution of reads as visualized in a boxplot of log counts. We observe `r sum(apply(log2(exprsDat+1),2,median)==0)` samples with zero median expression (`rpaste(names(which(apply(log2(exprsDat+1),2,median)==0)), collapse="; ")`) reflective of low read count to be filtered from subsequent analysis as low quality.

```{r}
boxplot(log2(exprsDat+1), ylab='log2(counts+1)', 
        xlab='samples', las=2, cex.axis = 0.25, 
        border = ifelse(apply(log2(exprsDat+1),2,median)==0,
                        'red','black'))
legend('topleft', pch=c('-'), legend = c('median > 0', 'median < 0'),
       col=c('black','red'), cex=0.5)

datatable(cbind(sampAnnot,
                lowCountFilter=apply(log2(exprsDat+1),2,median)>0))
```

```{r}
sampAnnotMedianFilter <- sampAnnot[apply(log2(exprsDat+1),2,median)>0,]
exprsDatMedian <- exprsDat[,apply(log2(exprsDat+1),2,median)>0]
```


## Obtain normalized log counts with vst 

[`DESeq2`](https://genomebiology.biomedcentral.com/articles/10.1186/s13059-014-0550-8) is used for variance stabilization transformation for count normalization. 

```{r}
dds <- DESeqDataSetFromMatrix(countData = round(exprsDatMedian), 
                              colData = sampAnnotMedianFilter, 
                              design=~Cell_Type)
vstDat <- assay(vst(dds))
vstDat <- vstDat[apply(vstDat,1,sd)>0,]
```

## Perform PCA analysis to evaluate sample clustering

We use PCA of the `vst` normalized data to evaluate sample clustering by cell type. We observe largely robust clustering by cell type, with an outlier CD8 and CD19 sample.

```{r}
pcs <- prcomp(t(vstDat),scale=T)
pca2d(pcs,
      group=sampAnnotMedianFilter$Cell_Type, legend="topleft")
title('PCA of vst counts after median filtering')
```

## Expression of marker genes

To further evaluate sample quality, we look at marker gene expression and its relationship to the sample groups observed in the PCA analysis.

```{r}
markerGenes <- c('CD4','CD8A','CD8B','ITGAM','CD19')

ht <- Heatmap(t(apply(vstDat[markerGenes,],1,scale)),
        clustering_distance_rows = 'pearson',
        clustering_distance_columns = 'pearson')

k5clust <- as.character(cutree(as.hclust(column_dend(ht)), k=5))
dfAnnot = data.frame(Cell_Type=sampAnnotMedianFilter[,'Cell_Type'],
                     cluster=k5clust)
row.names(dfAnnot) <- sampAnnotMedianFilter$Sample_id

HA <- HeatmapAnnotation(df = dfAnnot,
                        col=list(Cell_Type =
                                   c('CD11b'='orange','CD19'='blue',
                                     'CD4'='green','CD8'='yellow'),
                                 cluster = c('1'='yellow',
                                             '2'='green',
                                             '3'='blue',
                                             '4'='orange',
                                             '5'='red')))

Heatmap(t(apply(vstDat[markerGenes,],1,scale)),
        top_annotation = HA,
        clustering_distance_rows = 'pearson',
        clustering_distance_columns = 'pearson')

dfAnnot$Mislabel <- F
dfAnnot$Mislabel[dfAnnot$cluster=='5'] <- T
sampleClusters <- apply(table(k5clust,
                              sampAnnotMedianFilter$Cell_Type),2,
                        which.max)

for (c in names(sampleClusters)) {
  dfAnnot$Mislabel[dfAnnot$Cell_Type == c &
                     dfAnnot$cluster != sampleClusters[c]] <- T
}
```

Based on the clustering analysis analysis of marker genes, we identify potential mislabeling in the following samples:

```{r}
datatable(dfAnnot[dfAnnot$Mislabel,])
```

# Visualization of samples retained for analysis

```{r}
sampAnnotFilter <- sampAnnotMedianFilter[!dfAnnot$Mislabel,]
exprsDatFilter <- exprsDat[,rownames(sampAnnotFilter)]

save(sampAnnotFilter, file="~/Documents/Lei/sampAnnotFilter.rda")

dds <- DESeqDataSetFromMatrix(countData = round(exprsDatFilter), 
                              colData = sampAnnotFilter, 
                              design=~Cell_Type)
vstDat <- assay(vst(dds))
vstDat <- vstDat[apply(vstDat,1,sd)>0,]

pcs <- prcomp(t(vstDat),scale=T)
pca2d(pcs,
      group=sampAnnotFilter$Cell_Type, legend="topleft")
title('PCA of vst counts of filtered data')
```
```{r}
dfAnnot = data.frame(Cell_Type=sampAnnotFilter[,'Cell_Type'])
row.names(dfAnnot) <- sampAnnotFilter$Sample_id

HA <- HeatmapAnnotation(df = dfAnnot,
                        col=list(Cell_Type =
                                   c('CD11b'='orange','CD19'='blue',
                                     'CD4'='green','CD8'='yellow'),
                                 cluster = c('1'='green',
                                             '2'='yellow',
                                             '3'='orange',
                                             '4'='blue',
                                             '5'='red')))

Heatmap(t(apply(vstDat[markerGenes,],1,scale)),
        top_annotation = HA,
        clustering_distance_rows = 'pearson',
        clustering_distance_columns = 'pearson')
```

## Final sample numbers

```{r}
print("Treatment Samp Numbers")
table(sampAnnotFilter$Cell_Type,
      sampAnnotFilter$Treatment)
print("CD66b_Gr_Status Samp Numbers")
table(sampAnnotFilter$Cell_Type,
      sampAnnotFilter$CD66b_Gr_Status)
```

# Evaluation of batch effects

```{r}
batches <- read.csv('BatchAnnot.csv',sep=",", header=F)
batches$SampleID <- sprintf('%d_%03d_%s',
                            batches[,3], 
                            batches[,4],
                            batches[,5])
batches$SampleID[batches[,6]=='RLT'] <- 
  paste0(batches$SampleID[batches[,6]=='RLT'], '_RLT')
batches <- batches[!duplicated(batches$SampleID),]

row.names(batches) <- unique(batches$SampleID)




dfAnnot = data.frame(sampAnnotFilter[,c('Cell_Type',
                                                  'Treatment')],
                     batches[sampAnnotFilter$Sample_id,1:2])
row.names(dfAnnot) <- sampAnnotFilter$Sample_id

HA <- HeatmapAnnotation(df = dfAnnot,
                        col=list(Cell_Type =
                                   c('CD11b'='orange','CD19'='blue',
                                     'CD4'='green','CD8'='yellow')))

Heatmap(t(apply(vstDat[markerGenes,],1,scale)),
        top_annotation = HA,
        clustering_distance_rows = 'pearson',
        clustering_distance_columns = 'pearson')

pca2d(pcs,
      group=dfAnnot$V1, legend="topleft")
title('PCA of vst counts of filtered data by batch')

datCombat <- sva::ComBat(vstDat,
                         dfAnnot$V1,
                         mod=model.matrix(~dfAnnot$Cell_Type))

pcsCombat <- prcomp(t(datCombat),scale=T)
pca2d(pcsCombat,
      group=sampAnnotFilter$Cell_Type, legend="topright")
title('PCA of vst counts of filtered and batch corrected data')
```

# Differential expression and gene set analyses by cell type

For each cell type, we run differential expression analysis with [DESeq2](https://genomebiology.biomedcentral.com/articles/10.1186/s13059-014-0550-8), including batch as a covariate in our model. Estimated fold changes are shrunk with `apeglm` using `lfcShrink` to account for the variation in the samples in this dataset. Gene set statistics are run with [fgsea](https://www.biorxiv.org/content/10.1101/060012v3) with [MSigDb](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3106198/) v7.5.1 pathways annotated in [KEGG](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5210567/), REACTOME, [IMMUNESIGDB](http://dx.doi.org/10.1016/j.immuni.2015.12.006), and GO as well as a custom set of immune databases downloaded directly from these databases.

```{r}
HALLMARK <- gmtPathways('h.all.v7.5.1.symbols.gmt.txt')
KEGG <- gmtPathways('c2.cp.kegg.v7.5.1.symbols.gmt.txt')
REACTOME <- gmtPathways('c2.cp.reactome.v7.5.1.symbols.gmt.txt')
IMMUNESIG <- gmtPathways('c7.immunesigdb.v7.5.1.symbols.gmt.txt')
GOBP <- gmtPathways('c5.go.bp.v7.5.1.symbols.gmt.txt')
GOMF <- gmtPathways('c5.go.mf.v7.5.1.symbols.gmt.txt')

# obtain custom immune pathways (subset to human lists)
customImmune <- gmtPathways('genesetkeyunew.gmt')
customImmune <- sapply(customImmune,function(x){setdiff(x,"")})
customImmune <- customImmune[sapply(customImmune,
                                  function(x){all(x==toupper(x))})]

DEByCellType <- function(CellType) {
  sampByCell <- sampAnnotFilter[sampAnnotFilter$Cell_Type == CellType & !is.na(sampAnnotFilter$CD66b_Gr_Status),]
  sampIds <- rownames(sampByCell)
  
  ##CD66b_Gr_Status
  dds <- DESeqDataSetFromMatrix(countData =
                                  round(exprsDatFilter)[,colnames(exprsDatFilter) %in% sampIds], 
                              colData = sampByCell, 
                              design=~CD66b_Gr_Status+batch)
  dds <- DESeq(dds)
  res <- results(dds, name="CD66b_Gr_Status_Low_vs_High")
  res <- lfcShrink(dds, coef="CD66b_Gr_Status_Low_vs_High", 
                   res=res)
  res <- res[!is.na(res$padj),]
  
  stats <- res$log2FoldChange

  gsResults <- list(HALLMARK=fgsea(pathways=HALLMARK, 
                               stats=stats),
                    KEGG=fgsea(pathways=KEGG, 
                               stats=stats),
                    REACTOME=fgsea(pathways=REACTOME,
                                   stats=stats), 
                    IMMMUNESIG=fgsea(pathways=IMMUNESIG,
                                     stats=stats),
                    GOBP = fgsea(pathways=GOBP,
                                     stats=stats),
                    GOMF = fgsea(pathways=GOMF,
                                 stats=stats),
                    custom=fgsea(pathways=customImmune,
                                 stats=stats))
  
  return(list(res=res,gs=gsResults))
}
```

## CD4 cells

```{r}
CD4Res <- DEByCellType('CD4')
CD4DE <- CD4Res$res
datatable(data.frame(CD4DE))
EnhancedVolcano(CD4DE, lab=row.names(CD4DE), 
                x='log2FoldChange', y='padj', FCcutoff = 0.5,
                pCutoff = 0.05)

DEGenes <- row.names(CD4DE)[CD4DE$padj < 0.05 &
                              abs(CD4DE$log2FoldChange) > 0.5]
TopDE <- CD4DE[CD4DE$padj < 0.05 &
                              abs(CD4DE$log2FoldChange) > 0.5,]

write.csv(TopDE, "CD4_TopDE_CD66Gr.csv")

sampCD4 <- sampAnnotFilter[sampAnnotFilter$Cell_Type == 'CD4' & !is.na(sampAnnotFilter$CD66b_Gr_Status),]

dfAnnot = data.frame(Treatment=sampCD4[,'Treatment'],
                     Batches=batches[sampCD4$Sample_id,1:2],
                     CD66b_Gr_Status=sampCD4[,'CD66b_Gr_Status'])
```

```{r, fig.height=10}
HA <- HeatmapAnnotation(df = dfAnnot,
                        col=list(CD66b_Gr_Status =
                                   c('High'='blue',
                                     'Low'='red'), 
                                 Treatment =
                                   c('GVAX_CY'='blue',
                                     'GVAX_CY_NIVO'='red')))


Heatmap(t(apply(datCombat[DEGenes, rownames(sampCD4)],1,scale)),top_annotation = HA,
        clustering_distance_rows = 'pearson', 
        clustering_distance_columns = 'pearson',
        row_names_gp = gpar(fontsize = 6))
```

### HALLMARK statistics

```{r}
datatable(CD4Res$gs$HALLMARK)
fwrite((CD4Res$gs$HALLMARK), file="CD4_HALLMARK_CD66Gr.csv")

## Complete HALLMARK Barplot
CD4Res$gs$HALLMARK$adjPvalue <- ifelse(CD4Res$gs$HALLMARK$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD4Res$gs$HALLMARK, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Hallmark pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD4 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only HALLMARK Barplot
Sig_CD4ResHALLMARK <- CD4Res$gs$HALLMARK[padj < 0.05]
if (nrow(Sig_CD4ResHALLMARK)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD4ResHALLMARK, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) Hallmark pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD4 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}

```

### KEGG statistics

```{r}
datatable(CD4Res$gs$KEGG)
fwrite((CD4Res$gs$KEGG), file="CD4_KEGG_CD66Gr.csv")

## KEGG Barplot
CD4Res$gs$KEGG$adjPvalue <- ifelse(CD4Res$gs$KEGG$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD4Res$gs$KEGG, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=2)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="KEGG pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD4 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only KEGG Barplot
Sig_CD4ResKEGG <- CD4Res$gs$KEGG[padj < 0.05]
if (nrow(Sig_CD4ResKEGG)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD4ResKEGG, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) KEGG pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD4 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}

```

### REACTOME statistics

```{r}
datatable(CD4Res$gs$REACTOME)
fwrite((CD4Res$gs$REACTOME), file="CD4_REACTOME_CD66Gr.csv")

##REACTOME Barplot
CD4Res$gs$REACTOME$adjPvalue <- ifelse(CD4Res$gs$REACTOME$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD4Res$gs$REACTOME, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=1)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="REACTOME pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD4 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only REACTOME Barplot
Sig_CD4ResREACTOME <- CD4Res$gs$REACTOME[padj < 0.05]
if (nrow(Sig_CD4ResREACTOME)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD4ResREACTOME, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) REACTOME pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD4 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}
```

### GO BP statistics

```{r}
datatable(CD4Res$gs$GOBP)
fwrite((CD4Res$gs$GOBP), file="CD4_GOBP_CD66Gr.csv")

##GOBP Barplot
CD4Res$gs$GOBP$adjPvalue <- ifelse(CD4Res$gs$GOBP$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD4Res$gs$GOBP, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="GOBP pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD4 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only GOBP Barplot
Sig_CD4ResGOBP <- CD4Res$gs$GOBP[padj < 0.05]
if (nrow(Sig_CD4ResGOBP)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD4ResGOBP, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) GOBP pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD4 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}
```


### GO MF statistics

```{r}
datatable(CD4Res$gs$GOMF)
fwrite((CD4Res$gs$GOMF), file="CD4_GOMF_CD66Gr.csv")

##GOMF Barplot
CD4Res$gs$GOMF$adjPvalue <- ifelse(CD4Res$gs$GOMF$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD4Res$gs$GOMF, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="GOMF pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD4 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only GOMF Barplot
Sig_CD4ResGOMF <- CD4Res$gs$GOMF[padj < 0.05]
if (nrow(Sig_CD4ResGOMF)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD4ResGOMF, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) GOMF pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD4 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}
```


## CD8 cells

```{r}
CD8Res <- DEByCellType('CD8')
CD8DE <- CD8Res$res

datatable(data.frame(CD8DE))
EnhancedVolcano(CD8DE, lab=row.names(CD8DE), 
                x='log2FoldChange', y='padj', FCcutoff = 0.5,
                pCutoff = 0.05)

DEGenes <- row.names(CD8DE)[CD8DE$padj < 0.05 &
                              abs(CD8DE$log2FoldChange) > 0.5]
TopDE <- CD8DE[CD8DE$padj < 0.05 &
                              abs(CD8DE$log2FoldChange) > 0.5,]

write.csv(TopDE, "CD8_TopDE_CD66Gr.csv")

sampCD8 <- sampAnnotFilter[sampAnnotFilter$Cell_Type == 'CD8' & !is.na(sampAnnotFilter$CD66b_Gr_Status),]

dfAnnot = data.frame(Treatment=sampCD8[,'Treatment'],
                     Batches=batches[sampCD8$Sample_id,1:2],
                     CD66b_Gr_Status=sampCD8[,'CD66b_Gr_Status'])
```

```{r, fig.height=10}
HA <- HeatmapAnnotation(df = dfAnnot,
                        col=list(CD66b_Gr_Status =
                                   c('High'='blue',
                                     'Low'='red'), 
                                 Treatment =
                                   c('GVAX_CY'='blue',
                                     'GVAX_CY_NIVO'='red')))


Heatmap(t(apply(datCombat[DEGenes, rownames(sampCD8)],1,scale)),top_annotation = HA,
        clustering_distance_rows = 'pearson', 
        clustering_distance_columns = 'pearson',
        row_names_gp = gpar(fontsize = 6))
```

### HALLMARK statistics

```{r}
datatable(CD8Res$gs$HALLMARK)
fwrite((CD8Res$gs$HALLMARK), file="CD8_HALLMARK_CD66Gr.csv")

## Complete HALLMARK Barplot
CD8Res$gs$HALLMARK$adjPvalue <- ifelse(CD8Res$gs$HALLMARK$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD8Res$gs$HALLMARK, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Hallmark pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD8 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only HALLMARK Barplot
Sig_CD8ResHALLMARK <- CD8Res$gs$HALLMARK[padj < 0.05]
if (nrow(Sig_CD8ResHALLMARK)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD8ResHALLMARK, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) Hallmark pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD8 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}
```

### KEGG statistics

```{r}
datatable(CD8Res$gs$KEGG)
fwrite((CD8Res$gs$KEGG), file="CD8_KEGG_CD66Gr.csv")

## KEGG Barplot
CD8Res$gs$KEGG$adjPvalue <- ifelse(CD8Res$gs$KEGG$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD8Res$gs$KEGG, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=2)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="KEGG pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD8 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only KEGG Barplot
Sig_CD8ResKEGG <- CD8Res$gs$KEGG[padj < 0.05]
if (nrow(Sig_CD8ResKEGG)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD8ResKEGG, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) KEGG pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD8 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}

```

### REACTOME statistics

```{r}
datatable(CD8Res$gs$REACTOME)
fwrite((CD8Res$gs$REACTOME), file="CD8_REACTOME_CD66Gr.csv")

## REACTOME Barplot
CD8Res$gs$REACTOME$adjPvalue <- ifelse(CD8Res$gs$REACTOME$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD8Res$gs$REACTOME, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=2)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="REACTOME pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD8 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only REACTOME Barplot
Sig_CD8ResREACTOME <- CD8Res$gs$REACTOME[padj < 0.05]
if (nrow(Sig_CD8ResREACTOME)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD8ResREACTOME, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) REACTOME pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD8 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}

```

### GO BP statistics

```{r}
datatable(CD8Res$gs$GOBP)
fwrite((CD8Res$gs$GOBP), file="CD8_GOBP_CD66Gr.csv")

## GOBP Barplot
CD8Res$gs$GOBP$adjPvalue <- ifelse(CD8Res$gs$GOBP$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD8Res$gs$GOBP, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=2)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="GOBP pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD8 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only GOBP Barplot
Sig_CD8ResGOBP <- CD8Res$gs$GOBP[padj < 0.05]
if (nrow(Sig_CD8ResGOBP)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD8ResGOBP, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) GOBP pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD8 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}
```


### GO MF statistics

```{r}
datatable(CD8Res$gs$GOMF)
fwrite((CD8Res$gs$GOMF), file="CD8_GOMF_CD66Gr.csv")

## GOMF Barplot
CD8Res$gs$GOMF$adjPvalue <- ifelse(CD8Res$gs$GOMF$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD8Res$gs$GOMF, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=2)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="GOMF pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD8 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only GOMF Barplot
Sig_CD8ResGOMF <- CD8Res$gs$GOMF[padj < 0.05]
if (nrow(Sig_CD8ResGOMF)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD8ResGOMF, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) GOMF pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD8 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}
```


## CD19 cells

```{r}
CD19Res <- DEByCellType('CD19')
CD19DE <- CD19Res$res

datatable(data.frame(CD19DE))
EnhancedVolcano(CD19DE, lab=row.names(CD19DE), 
                x='log2FoldChange', y='padj', FCcutoff = 0.5,
                pCutoff = 0.05)

DEGenes <- row.names(CD19DE)[CD19DE$padj < 0.05 &
                              abs(CD19DE$log2FoldChange) > 0.5]
TopDE <- CD19DE[CD19DE$padj < 0.05 &
                              abs(CD19DE$log2FoldChange) > 0.5,]

write.csv(TopDE, "CD19_TopDE_CD66Gr.csv")

sampCD19 <- sampAnnotFilter[sampAnnotFilter$Cell_Type == 'CD19' & !is.na(sampAnnotFilter$CD66b_Gr_Status),]

dfAnnot = data.frame(Treatment=sampCD19[,'Treatment'],
                     Batches=batches[sampCD19$Sample_id,1:2],
                     CD66b_Gr_Status=sampCD19[,'CD66b_Gr_Status'])
```

```{r, fig.height=10}
HA <- HeatmapAnnotation(df = dfAnnot,
                        col=list(CD66b_Gr_Status =
                                   c('High'='blue',
                                     'Low'='red'), 
                                 Treatment =
                                   c('GVAX_CY'='blue',
                                     'GVAX_CY_NIVO'='red')))


Heatmap(t(apply(datCombat[DEGenes, rownames(sampCD19)],1,scale)),top_annotation = HA,
        clustering_distance_rows = 'pearson', 
        clustering_distance_columns = 'pearson',
        row_names_gp = gpar(fontsize = 6))
```

### HALLMARK statistics

```{r}
datatable(CD19Res$gs$HALLMARK)
fwrite((CD19Res$gs$HALLMARK), file="CD19_HALLMARK_CD66Gr.csv")

## HALLMARK Barplot
CD19Res$gs$HALLMARK$adjPvalue <- ifelse(CD19Res$gs$HALLMARK$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD19Res$gs$HALLMARK, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=2)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="HALLMARK pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD19 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only HALLMARK Barplot
Sig_CD19ResIMMMUNESIG <- CD19Res$gs$HALLMARK[padj < 0.05]
if (nrow(Sig_CD19ResIMMMUNESIG)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD19ResIMMMUNESIG, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) HALLMARK pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD19 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}
```

### KEGG statistics

```{r}
datatable(CD19Res$gs$KEGG)
fwrite((CD19Res$gs$KEGG), file="CD19_KEGG_CD66Gr.csv")

## KEGG Barplot
CD19Res$gs$KEGG$adjPvalue <- ifelse(CD19Res$gs$KEGG$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD19Res$gs$KEGG, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=2)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="KEGG pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD19 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only KEGG Barplot
Sig_CD19ResIMMMUNESIG <- CD19Res$gs$KEGG[padj < 0.05]
if (nrow(Sig_CD19ResIMMMUNESIG)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD19ResIMMMUNESIG, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) KEGG pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD19 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}
```

### REACTOME statistics

```{r}
datatable(CD19Res$gs$REACTOME)
fwrite((CD19Res$gs$REACTOME), file="CD19_REACTOME_CD66Gr.csv")

## REACTOME Barplot
CD19Res$gs$REACTOME$adjPvalue <- ifelse(CD19Res$gs$REACTOME$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD19Res$gs$REACTOME, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=2)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="REACTOME pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD19 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only REACTOME Barplot
Sig_CD19ResIMMMUNESIG <- CD19Res$gs$REACTOME[padj < 0.05]
if (nrow(Sig_CD19ResIMMMUNESIG)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD19ResIMMMUNESIG, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) REACTOME pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD19 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}
```

### GO BP statistics

```{r}
datatable(CD19Res$gs$GOBP)
fwrite((CD19Res$gs$GOBP), file="CD19_GOBP_CD66Gr.csv")

## GOBP Barplot
CD19Res$gs$GOBP$adjPvalue <- ifelse(CD19Res$gs$GOBP$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD19Res$gs$GOBP, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=2)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="GOBP pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD19 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only GOBP Barplot
Sig_CD19ResIMMMUNESIG <- CD19Res$gs$GOBP[padj < 0.05]
if (nrow(Sig_CD19ResIMMMUNESIG)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD19ResIMMMUNESIG, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) GOBP pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD19 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}
```


### GO MF statistics

```{r}
datatable(CD19Res$gs$GOMF)
fwrite((CD19Res$gs$GOMF), file="CD19_GOMF_CD66Gr.csv")

## GOMF Barplot
CD19Res$gs$GOMF$adjPvalue <- ifelse(CD19Res$gs$GOMF$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD19Res$gs$GOMF, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=2)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="GOMF pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD19 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only GOMF Barplot
Sig_CD19ResIMMMUNESIG <- CD19Res$gs$GOMF[padj < 0.05]
if (nrow(Sig_CD19ResIMMMUNESIG)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD19ResIMMMUNESIG, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) GOMF pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD19 Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}
```

## CD11b cells

```{r}
CD11bRes <- DEByCellType('CD11b')
CD11bDE <- CD11bRes$res
datatable(data.frame(CD11bDE))
EnhancedVolcano(CD11bDE, lab=row.names(CD11bDE), 
                x='log2FoldChange', y='padj',FCcutoff = 0.5,
                pCutoff = 0.05)

DEGenes <- row.names(CD11bDE)[CD11bDE$padj < 0.05 &
                              abs(CD11bDE$log2FoldChange) > 0.5]
TopDE <- CD11bDE[CD11bDE$padj < 0.05 &
                              abs(CD11bDE$log2FoldChange) > 0.5,]

write.csv(TopDE, "CD11b_TopDE_CD66Gr.csv")

sampCD11b <- sampAnnotFilter[sampAnnotFilter$Cell_Type == 'CD11b' & !is.na(sampAnnotFilter$CD66b_Gr_Status),]

dfAnnot = data.frame(Treatment=sampCD11b[,'Treatment'],
                     Batches=batches[sampCD11b$Sample_id,1:2],
                     CD66b_Gr_Status=sampCD11b[,'CD66b_Gr_Status'])
```

```{r, fig.height=10}
HA <- HeatmapAnnotation(df = dfAnnot,
                        col=list(CD66b_Gr_Status =
                                   c('High'='blue',
                                     'Low'='red'), 
                                 Treatment =
                                   c('GVAX_CY'='blue',
                                     'GVAX_CY_NIVO'='red')))


Heatmap(t(apply(datCombat[DEGenes, rownames(sampCD11b)],1,scale)),top_annotation = HA,
        clustering_distance_rows = 'pearson', 
        clustering_distance_columns = 'pearson',
        row_names_gp = gpar(fontsize = 6))
```

### HALLMARK statistics

```{r}
datatable(CD11bRes$gs$HALLMARK)
fwrite((CD11bRes$gs$HALLMARK), file="CD11b_HALLMARK_CD66Gr.csv")

## IMMMUNESIG Barplot
CD11bRes$gs$HALLMARK$adjPvalue <- ifelse(CD11bRes$gs$HALLMARK$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD11bRes$gs$HALLMARK, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=2)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="HALLMARK pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD11b Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only HALLMARK Barplot
Sig_CD11bResHALLMARK <- CD11bRes$gs$HALLMARK[padj < 0.05]
if (nrow(Sig_CD11bResHALLMARK)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD11bResHALLMARK, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) HALLMARK pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD11b Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}
```

### KEGG statistics

```{r}
datatable(CD11bRes$gs$KEGG)
fwrite((CD11bRes$gs$KEGG), file="CD11b_KEGG_CD66Gr.csv")

## KEGG Barplot
CD11bRes$gs$KEGG$adjPvalue <- ifelse(CD11bRes$gs$KEGG$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD11bRes$gs$KEGG, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=2)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="KEGG pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD11b Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only KEGG Barplot
Sig_CD11bResKEGG <- CD11bRes$gs$KEGG[padj < 0.05]
if (nrow(Sig_CD11bResKEGG)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD11bResKEGG, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) KEGG pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD11b Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}
```

### REACTOME statistics

```{r}
datatable(CD11bRes$gs$REACTOME)
fwrite((CD11bRes$gs$REACTOME), file="CD11b_REACTOME_CD66Gr.csv")

## REACTOME Barplot
CD11bRes$gs$REACTOME$adjPvalue <- ifelse(CD11bRes$gs$REACTOME$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD11bRes$gs$REACTOME, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=2)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="REACTOME pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD11b Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only REACTOME Barplot
Sig_CD11bResREACTOME <- CD11bRes$gs$REACTOME[padj < 0.05]
if (nrow(Sig_CD11bResREACTOME)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD11bResREACTOME, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) REACTOME pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD11b Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}
```

### GO BP statistics

```{r}
datatable(CD11bRes$gs$GOBP)
fwrite((CD11bRes$gs$GOBP), file="CD11b_GOBP_CD66Gr.csv")

## GOBP Barplot
CD11bRes$gs$GOBP$adjPvalue <- ifelse(CD11bRes$gs$GOBP$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD11bRes$gs$GOBP, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=2)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="GOBP pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD11b Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only GOBP Barplot
Sig_CD11bResGOBP <- CD11bRes$gs$GOBP[padj < 0.05]
if (nrow(Sig_CD11bResGOBP)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD11bResGOBP, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) GOBP pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD11b Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}
```


### GO MF statistics

```{r}
datatable(CD11bRes$gs$GOMF)
fwrite((CD11bRes$gs$GOMF), file="CD11b_GOMF_CD66Gr.csv")

## GOMF Barplot
CD11bRes$gs$GOMF$adjPvalue <- ifelse(CD11bRes$gs$GOMF$padj <= 0.05, "significant", "non-significant")
cols <- c("non-significant" = "grey", "significant" = "red")
ggplot(CD11bRes$gs$GOMF, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=2)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="GOMF pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD11b Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))

## Sig Only GOMF Barplot
Sig_CD11bResGOMF <- CD11bRes$gs$GOMF[padj < 0.05]
if (nrow(Sig_CD11bResGOMF)==0){ print("No Significant pathways with pAdj < 0.05"
)} else {
  ggplot(Sig_CD11bResGOMF, aes(reorder(pathway, NES), NES, fill = adjPvalue)) +
  geom_col() +
  scale_fill_manual(values = cols) +
  theme(axis.text.y = element_text(size=5)) +
  coord_flip() +
  labs(x="Pathway", y="Normalized Enrichment Score",
  title="Significant (pAdj<0.05) GOMF pathways Enrichment Score from CD66b_Gr_Status_Low_vs_High in CD11b Cells") + theme(plot.title.position = "plot", plot.title = element_text(size=8, face="bold"))
}
```


## Heatmap of Pathways of interest by sample
```{r}
##Extract Counts
Bulk_counts <- datCombat

#Gather Genes from msigdbr
hallmark_list = HALLMARK
TNFA_genes <- hallmark_list$HALLMARK_TNFA_SIGNALING_VIA_NFKB
EMT_genes <- hallmark_list$HALLMARK_EPITHELIAL_MESENCHYMAL_TRANSITION
Apop_genes <- hallmark_list$HALLMARK_APOPTOSIS
INFG_genes <- hallmark_list$HALLMARK_INTERFERON_GAMMA_RESPONSE
OXPhos_genes <- hallmark_list$HALLMARK_OXIDATIVE_PHOSPHORYLATION
INFA_genes <- hallmark_list$HALLMARK_INTERFERON_ALPHA_RESPONSE

Bulk_counts_TNFA <- as.data.frame(Bulk_counts[rownames(Bulk_counts) %in% TNFA_genes,])
Bulk_counts_EMT <- as.data.frame(Bulk_counts[rownames(Bulk_counts) %in% EMT_genes,])
Bulk_counts_Apop <- as.data.frame(Bulk_counts[rownames(Bulk_counts) %in% Apop_genes,])
Bulk_counts_INFG <- as.data.frame(Bulk_counts[rownames(Bulk_counts) %in% INFG_genes,])
Bulk_counts_OXPhos <- as.data.frame(Bulk_counts[rownames(Bulk_counts) %in% OXPhos_genes,])
Bulk_counts_INFA <- as.data.frame(Bulk_counts[rownames(Bulk_counts) %in% INFA_genes,])

##Heatmaps
pheatmap(Bulk_counts_TNFA, scale = "row", main="TNFA Heatmap", show_rownames = F)
pheatmap(Bulk_counts_EMT, scale = "row", main="EMT Heatmap", show_rownames = F)
pheatmap(Bulk_counts_Apop, scale = "row", main="Apop Heatmap", show_rownames = F)
pheatmap(Bulk_counts_INFG, scale = "row", main="INFG Heatmap", show_rownames = F)
pheatmap(Bulk_counts_OXPhos, scale = "row", main="OXPhos Heatmap", show_rownames = F)
pheatmap(Bulk_counts_INFA, scale = "row", main="INFA Heatmap", show_rownames = F)
```


# Cibersortx Results

Counts were extracted from the raw counts matrix (unsubset) and uploaded to [Cibersortx](https://cibersortx.stanford.edu). Imputation of cell fractions utilizing the LM22 CIBERSORT DEFAULT reference (signature matrix) to estimate T-cell presence based on gene expression. This was completed using [B-mode batch correction] (B-mode (bulk mode) batch correction removes technical differences between a signature matrix derived from bulk sorted reference profiles (e.g., bulk RNA-Seq or microarrays) and an input set of mixture samples. The technique can also be applied to signature matrices derived from scRNA-Seq platforms, provided that transcripts are measured analogously to bulk mixture expression profiles (e.g., full-length transcripts without UMIs profiled by SMART-Seq2)) and 1000 permutations for significance. 

## Basic Cibersort Results
### Load Data
```{r, }
CIBERSORTx_Res <- read.delim("CIBERSORTx_Job15_Adjusted.txt", row.names=1)
CIBERSORTx_Res_CellTypes <- CIBERSORTx_Res[,1:22]

#Add Treatment Arm to Results DF
CD66b_Gr_Status <- data.frame(sampAnnotFilter$CD66b_Gr_Status)
rownames(CD66b_Gr_Status) <- rownames(sampAnnotFilter)
colnames(CD66b_Gr_Status) <- "CD66b_Gr_Status"

CIBERSORTx_Res_CellTypes <- merge(CIBERSORTx_Res_CellTypes,CD66b_Gr_Status, by="row.names")
rownames(CIBERSORTx_Res_CellTypes) <- CIBERSORTx_Res_CellTypes$Row.names
CIBERSORTx_Res_CellTypes <- CIBERSORTx_Res_CellTypes[,-1]
```

### Heatmap
```{r}
#Display Results as Heatmap
CibersortAnnotation <- HeatmapAnnotation(Treatment= sampAnnotFilter[,5], OS= sampAnnotFilter[,8], Batch= sampAnnotFilter[,26], CellType=sampAnnotFilter[,4], CD66b_Gr_Status=sampAnnotFilter[,7], col=list(Treatment =
                                   c('GVAX_CY'='blue',
                                     'GVAX_CY_NIVO'='red')))

Heatmap(t(CIBERSORTx_Res_CellTypes[,1:22]), column_names_gp = gpar(fontsize = 4), row_names_gp = gpar(fontsize = 8), top_annotation = CibersortAnnotation,clustering_distance_rows = 'pearson', 
        clustering_distance_columns = 'pearson')
```

### Barplot
```{r}
#Display Results as Barplot
CIBERSORTx_Res_CellTypes %>% 
    mutate(Samples = rownames(.)) %>% # add row names as a variable
    reshape2::melt(.) %>% # melt to long format
    ggplot(., aes(x = Samples, y = value, fill = variable)) + 
        geom_bar(stat='identity')+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5)) + theme(legend.text = element_text(size=5)) +   theme(legend.key.size = unit(.3, 'cm')) + guides(fill=guide_legend(title="Cell Type")) + theme(legend.position = "bottom") + ggtitle("All Samples")

```

## Cibersort Results by CellType

### CD4
#### Subset Results
```{r}
CD4_cells <- rownames(sampAnnotFilter[sampAnnotFilter$Cell_Type=="CD4",])
CIBERSORTx_Res_CellTypes_CD4 <- CIBERSORTx_Res_CellTypes[rownames(CIBERSORTx_Res_CellTypes) %in% CD4_cells,]
```

#### Barplot
```{r}
CIBERSORTx_Res_CellTypes_CD4 %>% 
    mutate(Samples = rownames(.)) %>% # add row names as a variable
    reshape2::melt(.) %>% # melt to long format
    ggplot(., aes(x = Samples, y = value, fill = variable)) + 
        geom_bar(stat='identity')+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5)) + theme(legend.text = element_text(size=5)) +   theme(legend.key.size = unit(.3, 'cm')) + guides(fill=guide_legend(title="Cell Type")) + theme(legend.position = "bottom") + ggtitle("CD4 Samples")
```

#### Ttest by CD66b_Gr_Status
```{r}
datatable(CIBERSORTx_Res_CellTypes_CD4 %>% 
  gather(key = variable, value = value, -CD66b_Gr_Status) %>% 
  group_by(CD66b_Gr_Status, variable) %>% 
  summarise(value = list(value)) %>% 
  spread(CD66b_Gr_Status, value) %>% 
  group_by(variable) %>% 
  mutate(p_value = t.test(unlist(High), unlist(Low))$p.value,
         t_value = t.test(unlist(High), unlist(Low))$statistic))
```

### CD8
#### Subset Results
```{r}
CD8_cells <- rownames(sampAnnotFilter[sampAnnotFilter$Cell_Type=="CD8",])
CIBERSORTx_Res_CellTypes_CD8 <- CIBERSORTx_Res_CellTypes[rownames(CIBERSORTx_Res_CellTypes) %in% CD8_cells,]
```

#### Barplot
```{r}
CIBERSORTx_Res_CellTypes_CD8 %>% 
    mutate(Samples = rownames(.)) %>% # add row names as a variable
    reshape2::melt(.) %>% # melt to long format
    ggplot(., aes(x = Samples, y = value, fill = variable)) + 
        geom_bar(stat='identity')+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5)) + theme(legend.text = element_text(size=5)) +   theme(legend.key.size = unit(.3, 'cm')) + guides(fill=guide_legend(title="Cell Type")) + theme(legend.position = "bottom") + ggtitle("CD8 Samples")
```

#### Ttest by CD66b_Gr_Status
```{r}
datatable(CIBERSORTx_Res_CellTypes_CD8 %>% 
  gather(key = variable, value = value, -CD66b_Gr_Status) %>% 
  group_by(CD66b_Gr_Status, variable) %>% 
  summarise(value = list(value)) %>% 
  spread(CD66b_Gr_Status, value) %>% 
  group_by(variable) %>% 
  mutate(p_value = t.test(unlist(High), unlist(Low))$p.value,
         t_value = t.test(unlist(High), unlist(Low))$statistic))
```

### CD19
#### Subset Results
```{r}
CD19_cells <- rownames(sampAnnotFilter[sampAnnotFilter$Cell_Type=="CD19",])
CIBERSORTx_Res_CellTypes_CD19 <- CIBERSORTx_Res_CellTypes[rownames(CIBERSORTx_Res_CellTypes) %in% CD19_cells,]
```

#### Barplot
```{r}
CIBERSORTx_Res_CellTypes_CD19 %>% 
    mutate(Samples = rownames(.)) %>% # add row names as a variable
    reshape2::melt(.) %>% # melt to long format
    ggplot(., aes(x = Samples, y = value, fill = variable)) + 
        geom_bar(stat='identity')+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5)) + theme(legend.text = element_text(size=5)) +   theme(legend.key.size = unit(.3, 'cm')) + guides(fill=guide_legend(title="Cell Type")) + theme(legend.position = "bottom") + ggtitle("CD19 Samples")
```

#### Ttest by CD66b_Gr_Status
```{r}
datatable(CIBERSORTx_Res_CellTypes_CD19 %>% 
  gather(key = variable, value = value, -CD66b_Gr_Status) %>% 
  group_by(CD66b_Gr_Status, variable) %>% 
  summarise(value = list(value)) %>% 
  spread(CD66b_Gr_Status, value) %>% 
  group_by(variable) %>% 
  mutate(p_value = t.test(unlist(High), unlist(Low))$p.value,
         t_value = t.test(unlist(High), unlist(Low))$statistic))
```

### CD11b
#### Subset Results
```{r}
CD11b_cells <- rownames(sampAnnotFilter[sampAnnotFilter$Cell_Type=="CD11b",])
CIBERSORTx_Res_CellTypes_CD11b <- CIBERSORTx_Res_CellTypes[rownames(CIBERSORTx_Res_CellTypes) %in% CD11b_cells,]
```

#### Barplot
```{r}
CIBERSORTx_Res_CellTypes_CD11b %>% 
    mutate(Samples = rownames(.)) %>% # add row names as a variable
    reshape2::melt(.) %>% # melt to long format
    ggplot(., aes(x = Samples, y = value, fill = variable)) + 
        geom_bar(stat='identity')+ theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 5)) + theme(legend.text = element_text(size=5)) +   theme(legend.key.size = unit(.3, 'cm')) + guides(fill=guide_legend(title="Cell Type")) + theme(legend.position = "bottom") + ggtitle("CD11b Samples")
```

#### Ttest by CD66b_Gr_Status
```{r}
datatable(CIBERSORTx_Res_CellTypes_CD11b %>% 
  gather(key = variable, value = value, -CD66b_Gr_Status) %>% 
  group_by(CD66b_Gr_Status, variable) %>% 
  summarise(value = list(value)) %>% 
  spread(CD66b_Gr_Status, value) %>% 
  group_by(variable) %>% 
  mutate(p_value = t.test(unlist(High), unlist(High))$p.value,
         t_value = t.test(unlist(Low), unlist(Low))$statistic))
```

# Neoantigen Results
ADD DESCRIPTION OF NEOANTIGEN PREDICTIONS METHOD

## Load Neoantigen SampInfo
```{r}
Neoantigen_sampInfo = read.csv('SampInfo_TMB_fromMAF.csv')
Neoantigen_sampInfo = Neoantigen_sampInfo[Neoantigen_sampInfo$Treatment %in% c("GVAX","Nivo") & Neoantigen_sampInfo$CD66b_Gr_Status %in% c("High","Low"),]
```

## Load Neoantigen Predictions
```{r}
Neoantigen <- read_excel("Neoantigen_prediction 19s_(DNA-seq).xlsx", sheet = "HLA_A.19s.neoantigen.countstat")
##Separate SampNames
Neoantigen$X <- gsub(".*-VS-","",Neoantigen$Sample)
```

## Boxplots of mutationsite_valid
```{r}
Neoantigen_plotdf <- merge(Neoantigen_sampInfo,Neoantigen,by="X")

ggplot(Neoantigen_plotdf, aes(x=CD66b_Gr_Status, y=Total_mutationsite_valid)) + geom_boxplot() + geom_point(aes(col=X)) + stat_compare_means(method = "t.test") + ggtitle("Boxplot of Total_mutationsite_valid by CD66b_Gr_Status")
ggplot(Neoantigen_plotdf, aes(x=CD66b_Gr_Status, y=netMHCpan_total_mutationsite_valid)) + geom_boxplot() + geom_point(aes(col=X)) + stat_compare_means(method = "t.test") + ggtitle("Boxplot of netMHCpan_total_mutationsite_valid by CD66b_Gr_Status")
```

# TCR Results
Volume and Clones were extracted from the repExplore function of [IMMUNARCH]. Volume=calculates the number of unique clonotypes in the input data. Clones=returns the number of clones (i.e., cells) per input repertoire.

## Load Results from TCR Script
```{r}
load("tcrcloneinfo.rda")
```

## Differential expression w TCR counts by cell type

For each cell type, we run differential expression analysis with [DESeq2](https://genomebiology.biomedcentral.com/articles/10.1186/s13059-014-0550-8), including batch as a covariate in our model. Estimated fold changes are shrunk with `apeglm` using `lfcShrink` to account for the variation in the samples in this dataset.

```{r}
DETCRByCellType <- function(CellType, TCR) {
  sampByCell <- sampAnnotFilter[sampAnnotFilter$Cell_Type == CellType & !is.na(sampAnnotFilter$CD66b_Gr_Status),]
  sampIds <- rownames(sampByCell)
  
  ##Extract CountData
  countData = round(exprsDatFilter)[,colnames(exprsDatFilter) %in% sampIds]
  ##Extract sampleNames
  countNames = gsub("X.._12","12",colnames(countData))
  countNames = gsub("X","",countNames)
  countNames = gsub("_",".",countNames)
  ##Order TCR by countNames
  TCR <- as.data.frame(TCR)
  rownames(TCR) <- TCR$SampleNames
  TCR <- TCR[countNames,]
  TCR <- t(TCR)
  Volume <- as.numeric(TCR[2,])
  Clones <- as.numeric(TCR[3,])
  ##Append to Counts
  countData <- rbind(Volume,Clones,countData)
  
  ##Treatment OS
  dds <- DESeqDataSetFromMatrix(countData = countData, 
                              colData = sampByCell, 
                              design=~CD66b_Gr_Status+batch)
  dds <- DESeq(dds)
  res <- results(dds, name="CD66b_Gr_Status_Low_vs_High")
  res <- lfcShrink(dds, coef="CD66b_Gr_Status_Low_vs_High", 
                   res=res)
  res <- res[!is.na(res$padj),]
  
  return(res)
}
```

##CD8 Cells
```{r}
CD8TCR_DERes <- DETCRByCellType("CD8", cd8TCR)

datatable(data.frame(CD8TCR_DERes[1:2,]))
```

##CD4 Cells
```{r}
CD4TCR_DERes <- DETCRByCellType("CD4", cd4TCR)

datatable(data.frame(CD4TCR_DERes[1:2,]))
```
