---
title: "TCRseq_analysis - beta chain"
author: "Joe and Janelle"
date: "6/21/2022"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "~/OneDrive - Johns Hopkins/JHU_PDAC_Collaboration/Analysis_HTMLs/") })
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages
```{r}
library(immunarch)
library(readxl)
library(ggpubr)
library(ggplot2)
sessionInfo()
```

## Load MIXCR Results -- https://immunarch.com/articles/web_only/load_mixcr.html
```{r}
#Replace with the path to the folder with your processed MiXCR data.
mixcr_path = "~/OneDrive - Johns Hopkins/JHU_PDAC_Collaboration/MIXCR_Res"

##CD8
CD8_files = intersect(list.files(mixcr_path, recursive = T,full.names = T, pattern =".CD8"),list.files(mixcr_path, recursive = T,full.names = T, pattern =".clonotypes.TRB"))
CD8_immdata_mixcr <- repLoad(CD8_files)
###Save
saveRDS(CD8_immdata_mixcr, file="CD8_immdata_trb.rds")

##CD4
CD4_files = intersect(list.files(mixcr_path, recursive = T,full.names = T, pattern =".CD4"),list.files(mixcr_path, recursive = T,full.names = T, pattern =".clonotypes.TRB"))
CD4_immdata_mixcr <- repLoad(CD4_files)
###Save
saveRDS(CD4_immdata_mixcr, file="CD4_immdata_trb.rds")

```

## Load SampInfo
```{r}
load("sampAnnotFilter.Rda")
sampAnnot <- sampAnnotFilter
sampAnnot$Treatment[sampAnnot$Treatment == 'GVAX_CY_Nivo'] <- 
  'GVAX_CY_NIVO'
rownames(sampAnnot) <- make.names(sampAnnot$Sample_id)
sampAnnot$SampleNames <- paste(sampAnnot$PatientID, sampAnnot$Cell_Type,sep="_")
sampAnnot$SampleNames <- gsub("_",".", sampAnnot$SampleNames)

#CD8
##Extract SampleNames from MIXCR$meta
filenames <- CD8_immdata_mixcr$meta$Sample
#Remove batch1 additions
samplenames = gsub("Plus.+","",filenames)
#Remove batch2 additions
samplenames = gsub("_.+","",samplenames)
samplenames = gsub("B.","",samplenames)
##Adjust for the two sets of 12.020 samples
samplenames[8] = "13.020.CD8"
##SaveFinalSampleNames
samplenamesCD8 <- samplenames

CD8_immdata_mixcr$meta$SampleNames <- samplenames

##Extract Patient Names
samplenames = gsub(".CD8","",samplenames)
CD8_immdata_mixcr$meta$PatientID <- samplenames

##Subset to be samples in sampAnnot
CD8_immdata_mixcr <- repFilter(CD8_immdata_mixcr, "by.meta", list(SampleNames = include(sampAnnot$SampleNames)))

##Merge SampAnnot w/ Meta
CD8_immdata_mixcr$meta <- merge(CD8_immdata_mixcr$meta, sampAnnot, by="SampleNames")
#colnames(CD8_immdata_mixcr$meta)[1] <- "Sample" 
#colnames(CD8_immdata_mixcr$meta)[2] <- "FileNames"
#names(CD8_immdata_mixcr$data) <- CD8_immdata_mixcr$meta$Sample

##Save
saveRDS(CD8_immdata_mixcr, file="CD8_immdata_trb_wMETA.rds")

###-----------------------------------------###

#CD4
##Extract SampleNames from MIXCR$meta
filenames <- CD4_immdata_mixcr$meta$Sample
#Remove batch1 additions
samplenames = gsub("Plus.+","",filenames)
#Remove batch2 additions
samplenames = gsub("_.+","",samplenames)
samplenames = gsub("B.","",samplenames)
##SaveFinalSampleNames
samplenamesCD4 <- samplenames

CD4_immdata_mixcr$meta$SampleNames <- samplenames

##Extract Patient Names
samplenames = gsub(".CD4","",samplenames)
CD4_immdata_mixcr$meta$PatientID <- samplenames

##Subset to be samples in sampAnnot
CD4_immdata_mixcr <- repFilter(CD4_immdata_mixcr, "by.meta", list(SampleNames = include(sampAnnot$SampleNames)))

##Merge SampAnnot w/ Meta
CD4_immdata_mixcr$meta <- merge(CD4_immdata_mixcr$meta, sampAnnot, by="SampleNames")
#colnames(CD4_immdata_mixcr$meta)[1] <- "Sample" 
#colnames(CD4_immdata_mixcr$meta)[2] <- "FileNames"
#names(CD4_immdata_mixcr$data) <- CD4_immdata_mixcr$meta$Sample

##Save
saveRDS(CD4_immdata_mixcr, file="CD4_immdata_trb_wMETA.rds")

```


## CD8 Basic Immunarch Plots -- pdf saved CD8_immdata_mixcr_trb_sampAnnotFilter.pdf
```{r, include=FALSE}
pdf(file = paste0("CD8_immdata_mixcr_trb_sampAnnotFilter_ttest.pdf"))

##assign data
dat= get("CD8_immdata_mixcr")

##Clones by Annotation
print(tryCatch({vis(repExplore(dat$data, .method = "clones"), .by = c("Treatment"), .meta = dat$meta, .test=F) + ggtitle("Number of CD8 T cells by Treatment") + labs(x= "Treatment") + stat_compare_means(comparisons = list(c("GVAX_CY", "GVAX_CY_NIVO")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clones"), .by = c("OS"), .meta = dat$meta, .test=F) + ggtitle("Number of CD8 T cells by OS") + labs(x= "OS")+ stat_compare_means(comparisons = list(c("<2y", ">2y")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clones"), .by = c("CD66b_Gr_Status"), .meta = dat$meta, .test=F) + ggtitle("Number of CD8 T cells by CD66b_Gr_Status") + labs(x= "CD66b_Gr_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clones"), .by = c("Treg_Status"), .meta = dat$meta, .test=F) + ggtitle("Number of CD8 T cells by Treg_Status")  + labs(x= "Treg_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clones"), .by = c("IL17_Signaling_Status"), .meta = dat$meta, .test=F) + ggtitle("Number of CD8 T cells by IL17_Signaling_Status")  + labs(x= "IL17_Signaling_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clones"), .by = c("CD4_Th17_Status"), .meta = dat$meta, .test=F) + ggtitle("Number of CD8 T cells by CD4_Th17_Status")  + labs(x= "CD4_Th17_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clones"), .by = c("CD8_CD137_Status"), .meta = dat$meta, .test=F) + ggtitle("Number of CD8 T cells by CD8_CD137_Status")  + labs(x= "CD8_CD137_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(vis(repExplore(dat$data), .method = "volume")+ scale_x_discrete(labels=dat$meta$Publication_id))
print(vis(repExplore(dat$data))+ scale_x_discrete(labels=dat$meta$Publication_id))


##Clonal Proportion plots
print(vis(repClonality(dat$data, .method = "clonal.prop")) + scale_x_discrete(labels=dat$meta$Publication_id))
print(tryCatch({vis(repExplore(dat$data, .method = "clonal.prop"), .by = c("Treatment"), .meta = dat$meta, .test=F) + labs(x= "Treatment") + stat_compare_means(comparisons = list(c("GVAX_CY", "GVAX_CY_NIVO")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clonal.prop"), .by = c("OS"), .meta = dat$meta, .test=F) + labs(x= "OS")+ stat_compare_means(comparisons = list(c("<2y", ">2y")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clonal.prop"), .by = c("CD66b_Gr_Status"), .meta = dat$meta, .test=F) + labs(x= "CD66b_Gr_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clonal.prop"), .by = c("Treg_Status"), .meta = dat$meta, .test=F) + labs(x= "Treg_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clonal.prop"), .by = c("IL17_Signaling_Status"), .meta = dat$meta, .test=F) + labs(x= "IL17_Signaling_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clonal.prop"), .by = c("CD4_Th17_Status"), .meta = dat$meta, .test=F) + labs(x= "CD4_Th17_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clonal.prop"), .by = c("CD8_CD137_Status"), .meta = dat$meta, .test=F) + labs(x= "CD8_CD137_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))

##Top clonal proportion
print(tryCatch({vis(repClonality(dat$data, .method = "top", .head = c(10, 100, 1000, 3000, 10000)), .by = "Treatment", .meta = dat$meta)}, error=function(err){}))
print(tryCatch({vis(repClonality(dat$data, .method = "top", .head = c(10, 100, 1000, 3000, 10000)), .by = "OS", .meta = dat$meta)}, error=function(err){}))
print(tryCatch({vis(repClonality(dat$data, .method = "top", .head = c(10, 100, 1000, 3000, 10000)), .by = "CD66b_Gr_Status", .meta = dat$meta)}, error= function(err){}))
print(tryCatch({vis(repClonality(dat$data, .method = "top", .head = c(10, 100, 1000, 3000, 10000)), .by = "Treg_Status", .meta = dat$meta)}, error= function(err){}))
print(tryCatch({vis(repClonality(dat$data, .method = "top", .head = c(10, 100, 1000, 3000, 10000)), .by = "IL17_Signaling_Status", .meta = dat$meta)}, error= function(err){}))
print(tryCatch({vis(repClonality(dat$data, .method = "top", .head = c(10, 100, 1000, 3000, 10000)), .by = "CD4_Th17_Status", .meta = dat$meta)}, error= function(err){}))
print(tryCatch({vis(repClonality(dat$data, .method = "top", .head = c(10, 100, 1000, 3000, 10000)), .by = "CD8_CD137_Status", .meta = dat$meta)}, error= function(err){}))

##Repertoire overlap
print(tryCatch({vis(repOverlap(dat$data, .method = "public"), .verbose = F) + labs(title="Public") + scale_x_discrete(labels=dat$meta$Publication_id) + scale_y_discrete(labels=rev(dat$meta$Publication_id))}, error=function(err){}))
print(tryCatch({vis(repOverlap(dat$data, .method = "morisita", .verbose = F), .signif.digits = 2) + labs(title="Morisita") + scale_x_discrete(labels=dat$meta$Publication_id) + scale_y_discrete(labels=rev(dat$meta$Publication_id))}, error=function(err){}))

##Repertoire Diversity
# Chao1 diversity measure
print(vis(repDiversity(dat$data, "chao1")) +scale_x_discrete(labels=dat$meta$Publication_id))
# Hill numbers
print(vis(repDiversity(dat$data, "hill"))+scale_fill_discrete(labels=dat$meta$Publication_id))
# D50
print(vis(repDiversity(dat$data, "d50"))+scale_x_discrete(labels=dat$meta$Publication_id))
# Ecological diversity measure
print(vis(repDiversity(dat$data, "div"))+scale_x_discrete(labels=dat$meta$Publication_id))

dev.off()
```
## CD4 Basic Immunarch Plots -- pdf saved CD4_immdata_mixcr_trb_sampAnnotFilter.pdf
```{r, include=FALSE}
pdf(file = paste0("CD4_immdata_mixcr_trb_sampAnnotFilter_ttest.pdf"))

##assign data
dat= get("CD4_immdata_mixcr")

##Clones by Annotation
print(tryCatch({vis(repExplore(dat$data, .method = "clones"), .by = c("Treatment"), .meta = dat$meta, .test=F) + ggtitle("Number of CD4 T cells by Treatment") + labs(x= "Treatment") + stat_compare_means(comparisons = list(c("GVAX_CY", "GVAX_CY_NIVO")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clones"), .by = c("OS"), .meta = dat$meta, .test=F) + ggtitle("Number of CD4 T cells by OS") + labs(x= "OS")+ stat_compare_means(comparisons = list(c("<2y", ">2y")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clones"), .by = c("CD66b_Gr_Status"), .meta = dat$meta, .test=F) + ggtitle("Number of CD4 T cells by CD66b_Gr_Status") + labs(x= "CD66b_Gr_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clones"), .by = c("Treg_Status"), .meta = dat$meta, .test=F) + ggtitle("Number of CD4 T cells by Treg_Status")  + labs(x= "Treg_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clones"), .by = c("IL17_Signaling_Status"), .meta = dat$meta, .test=F) + ggtitle("Number of CD4 T cells by IL17_Signaling_Status")  + labs(x= "IL17_Signaling_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clones"), .by = c("CD4_Th17_Status"), .meta = dat$meta, .test=F) + ggtitle("Number of CD4 T cells by CD4_Th17_Status")  + labs(x= "CD4_Th17_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clones"), .by = c("CD8_CD137_Status"), .meta = dat$meta, .test=F) + ggtitle("Number of CD4 T cells by CD8_CD137_Status")  + labs(x= "CD8_CD137_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(vis(repExplore(dat$data), .method = "volume")+ scale_x_discrete(labels=dat$meta$Publication_id))
print(vis(repExplore(dat$data))+ scale_x_discrete(labels=dat$meta$Publication_id))


##Clonal Proportion plots
print(vis(repClonality(dat$data, .method = "clonal.prop")) + scale_x_discrete(labels=dat$meta$Publication_id))
print(tryCatch({vis(repExplore(dat$data, .method = "clonal.prop"), .by = c("Treatment"), .meta = dat$meta, .test=F) + labs(x= "Treatment") + stat_compare_means(comparisons = list(c("GVAX_CY", "GVAX_CY_NIVO")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clonal.prop"), .by = c("OS"), .meta = dat$meta, .test=F) + labs(x= "OS")+ stat_compare_means(comparisons = list(c("<2y", ">2y")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clonal.prop"), .by = c("CD66b_Gr_Status"), .meta = dat$meta, .test=F) + labs(x= "CD66b_Gr_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clonal.prop"), .by = c("Treg_Status"), .meta = dat$meta, .test=F) + labs(x= "Treg_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clonal.prop"), .by = c("IL17_Signaling_Status"), .meta = dat$meta, .test=F) + labs(x= "IL17_Signaling_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clonal.prop"), .by = c("CD4_Th17_Status"), .meta = dat$meta, .test=F) + labs(x= "CD4_Th17_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))
print(tryCatch({vis(repExplore(dat$data, .method = "clonal.prop"), .by = c("CD8_CD137_Status"), .meta = dat$meta, .test=F) + labs(x= "CD8_CD137_Status") + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")}, error= function(err){}))

##Top clonal proportion
print(tryCatch({vis(repClonality(dat$data, .method = "top", .head = c(10, 100, 1000, 3000, 10000)), .by = "Treatment", .meta = dat$meta)}, error=function(err){}))
print(tryCatch({vis(repClonality(dat$data, .method = "top", .head = c(10, 100, 1000, 3000, 10000)), .by = "OS", .meta = dat$meta)}, error=function(err){}))
print(tryCatch({vis(repClonality(dat$data, .method = "top", .head = c(10, 100, 1000, 3000, 10000)), .by = "CD66b_Gr_Status", .meta = dat$meta)}, error= function(err){}))
print(tryCatch({vis(repClonality(dat$data, .method = "top", .head = c(10, 100, 1000, 3000, 10000)), .by = "Treg_Status", .meta = dat$meta)}, error= function(err){}))
print(tryCatch({vis(repClonality(dat$data, .method = "top", .head = c(10, 100, 1000, 3000, 10000)), .by = "IL17_Signaling_Status", .meta = dat$meta)}, error= function(err){}))
print(tryCatch({vis(repClonality(dat$data, .method = "top", .head = c(10, 100, 1000, 3000, 10000)), .by = "CD4_Th17_Status", .meta = dat$meta)}, error= function(err){}))
print(tryCatch({vis(repClonality(dat$data, .method = "top", .head = c(10, 100, 1000, 3000, 10000)), .by = "CD8_CD137_Status", .meta = dat$meta)}, error= function(err){}))

##Repertoire overlap
print(tryCatch({vis(repOverlap(dat$data, .method = "public"), .verbose = F) + labs(title="Public") + scale_x_discrete(labels=dat$meta$Publication_id) + scale_y_discrete(labels=rev(dat$meta$Publication_id))}, error=function(err){}))
print(tryCatch({vis(repOverlap(dat$data, .method = "morisita", .verbose = F), .signif.digits = 2) + labs(title="Morisita") + scale_x_discrete(labels=dat$meta$Publication_id) + scale_y_discrete(labels=rev(dat$meta$Publication_id))}, error=function(err){}))

##Repertoire Diversity
# Chao1 diversity measure
print(vis(repDiversity(dat$data, "chao1")) +scale_x_discrete(labels=dat$meta$Publication_id))
# Hill numbers
print(vis(repDiversity(dat$data, "hill"))+scale_fill_discrete(labels=dat$meta$Publication_id))
# D50
print(vis(repDiversity(dat$data, "d50"))+scale_x_discrete(labels=dat$meta$Publication_id))
# Ecological diversity measure
print(vis(repDiversity(dat$data, "div"))+scale_x_discrete(labels=dat$meta$Publication_id))

dev.off()

```

## CD4 Clonality -- JMM July 14, 2022 -- pdf saved CD4_immdata_mixcr_trb_clonality.pdf
```{r}
##CD4s
####Clonality: Normalized Shannon's Entropy (Shannon's equitability- measure of evenness)
#Not so reliable when there are few (<100) cells
#between 0 and 1, lower value indicates more evenness (more polyclonal), higher value indicates less evenness (1 is monoclonal)
div_div <- repDiversity(coding(CD4_immdata_mixcr$data), "div", .q = 1, .col ="aa") #.q = 1 gives you exp shannon
div_div$clonotypes <- 0
#convert from true diversity back to Shannon entropy
div_div$Value <- log(div_div$Value)
#Get number of clonotypes per sample to calculate normalized Shannon entropy
for (i in 1:length(CD4_immdata_mixcr[[1]])){
  sample <- names(CD4_immdata_mixcr[[1]][i])
  test <- data.frame(coding(CD4_immdata_mixcr[[1]][i]))
  clonotypes <- length(unique(test[,4]))##make it based on amino acid clonotypes
  div_div$clonotypes[i] <- clonotypes
}

#div_div$clonotypes <- div_div$clonotypes + 0.1 #to avoid 0 values
#Calculate normalized Shannon entropy 
div_div$lnClonotypes <- log(div_div$clonotypes)
div_div$shannon_norm <- div_div$Value/div_div$lnClonotypes
div_div$Value <- 1-div_div$shannon_norm

##Assign clonality to metadata
clonalityDF <- div_div[,c(1:2)]
colnames(clonalityDF) <- c("Sample", "clonality")
CD4_immdata_mixcr$meta <- merge(CD4_immdata_mixcr$meta, clonalityDF, by = "Sample")

##Plots
pdf(file="CD4_immdata_mixcr_trb_clonality.pdf")

vis(div_div) + scale_x_discrete(labels=CD4_immdata_mixcr$meta$Publication_id)+ggtitle("Clonality")+ylab("Clonality")+ylim(0,0.5)

vis(div_div, .by = "OS", .meta = CD4_immdata_mixcr$meta, .test = T) + ggtitle("Clonality by OS")+ylab("Clonality")+ylim(0,0.5)
#on plot p=0.8
wilcox.test(CD4_immdata_mixcr$meta$clonality ~ CD4_immdata_mixcr$meta$OS, exact = F)
#p=0.7595
t.test(CD4_immdata_mixcr$meta$clonality ~ CD4_immdata_mixcr$meta$OS, exact = F)
#0.6728

vis(div_div, .by = "OS", .meta = CD4_immdata_mixcr$meta, .test = F) + ggtitle("CD4 T cell Clonality by OS")+ylab("Clonality")+ ylim(0,0.5) + stat_compare_means(comparisons = list(c("<2y", ">2y")), method = "t.test")


vis(div_div, .by = "Treatment", .meta = CD4_immdata_mixcr$meta, .test = T) + ggtitle("Clonality by Treatment")+ylab("Clonality")+ylim(0,0.5)
#p=0.09
wilcox.test(CD4_immdata_mixcr$meta$clonality ~ CD4_immdata_mixcr$meta$Treatment, exact = F)
#p=0.09
t.test(CD4_immdata_mixcr$meta$clonality ~ CD4_immdata_mixcr$meta$Treatment, exact = F)
#0.3404
vis(div_div, .by = "Treatment", .meta = CD4_immdata_mixcr$meta, .test = F) + ggtitle("CD4 T cell Clonality by Treatment")+ylab("Clonality")+ylim(0,0.5)+stat_compare_means(comparisons = list(c("GVAX_CY", "GVAX_CY_NIVO")), method = "t.test")

vis(div_div, .by = "CD66b_Gr_Status", .meta = CD4_immdata_mixcr$meta) + ggtitle("Clonality by CD66b_Gr_Status")+ylab("Clonality")+ylim(0,0.5)
#p=0.07
wilcox.test(CD4_immdata_mixcr$meta$clonality ~ CD4_immdata_mixcr$meta$CD66b_Gr_Status, exact = F)
#p=0.0726
t.test(CD4_immdata_mixcr$meta$clonality ~ CD4_immdata_mixcr$meta$CD66b_Gr_Status, exact = F)
#0.06187
vis(div_div, .by = "CD66b_Gr_Status", .meta = CD4_immdata_mixcr$meta, .test=FALSE) + ggtitle("CD4 T cell Clonality by CD66b_Gr_Status")+ylab("Clonality")+ylim(0,0.5) + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")

vis(div_div, .by = "CD8_CD137_Status", .meta = CD4_immdata_mixcr$meta) + ggtitle("Clonality by CD8_CD137_Status")+ylab("Clonality")+ylim(0,0.5)
#p=0.6
wilcox.test(CD4_immdata_mixcr$meta$clonality ~ CD4_immdata_mixcr$meta$CD8_CD137_Status, exact = F)
#p=0.5959
t.test(CD4_immdata_mixcr$meta$clonality ~ CD4_immdata_mixcr$meta$CD8_CD137_Status, exact = F)
#0.4134
vis(div_div, .by = "CD8_CD137_Status", .meta = CD4_immdata_mixcr$meta, .test=F) + ggtitle("CD4 T cell Clonality by CD8_CD137_Status")+ylab("Clonality")+ylim(0,0.5) + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")

vis(div_div, .by = "CD4_Th17_Status", .meta = CD4_immdata_mixcr$meta) + ggtitle("Clonality by CD4_Th17_Status")+ylab("Clonality")+ylim(0,0.5)
#p=0.05
wilcox.test(CD4_immdata_mixcr$meta$clonality ~ CD4_immdata_mixcr$meta$CD4_Th17_Status, exact = F)
#p=0.05523
t.test(CD4_immdata_mixcr$meta$clonality ~ CD4_immdata_mixcr$meta$CD4_Th17_Status, exact = F)
#0.05408
vis(div_div, .by = "CD4_Th17_Status", .meta = CD4_immdata_mixcr$meta, .test = F) + ggtitle("CD4 T cell Clonality by CD4_Th17_Status")+ylab("Clonality")+ylim(0,0.5) + stat_compare_means(comparisons = list(c("High", "Low")), method = "t.test")

vis(div_div, .by = "Treg_Status", .meta = CD4_immdata_mixcr$meta, .test = F) + ggtitle("CD4 T cell Clonality by Treg_Status")+ylab("Clonality")+ylim(0,0.65) + stat_compare_means(comparisons = list(c("High","Low")),method = "t.test")

vis(div_div, .by = "IL17_Signaling_Status", .meta = CD4_immdata_mixcr$meta, .test = F) + ggtitle("CD4 T cell Clonality by IL17_Signaling_Status")+ylab("Clonality")+ylim(0,0.65) + stat_compare_means(comparisons = list(c("High","Low")),method = "t.test")

dev.off()
```

## CD8 Clonality -- JMM July 14, 2022 -- pdf saved CD8_immdata_mixcr_trb_clonality.pdf
```{r}
###CD8s
####Clonality: Normalized Shannon's Entropy (Shannon's equitability- measure of evenness)
#Not so reliable when there are few (<100) cells
#between 0 and 1, lower value indicates more evenness (more polyclonal), higher value indicates less evenness (1 is monoclonal)
div_div <- repDiversity(coding(CD8_immdata_mixcr$data), "div", .q = 1, .col ="aa") #.q = 1 gives you exp shannon
div_div$clonotypes <- 0
#convert from true diversity back to Shannon ent
div_div$Value <- log(div_div$Value)
#Get number of clonotypes per sample to calculate normalized Shannon entropy
for (i in 1:length(CD8_immdata_mixcr[[1]])){
  sample <- names(CD8_immdata_mixcr[[1]][i])
  test <- data.frame(coding(CD8_immdata_mixcr[[1]][i]))
  clonotypes <- length(unique(test[,4]))
  div_div$clonotypes[i] <- clonotypes
}

#div_div$clonotypes <- div_div$clonotypes + 0.1 #to avoid 0 values
#Calculate normalized Shannon entropy 
div_div$lnClonotypes <- log(div_div$clonotypes)
div_div$shannon_norm <- div_div$Value/div_div$lnClonotypes
div_div$Value <- 1-div_div$shannon_norm

##Assign clonality to metadata
clonalityDF <- div_div[,c(1:2)]
colnames(clonalityDF) <- c("Sample", "clonality")
CD8_immdata_mixcr$meta <- merge(CD8_immdata_mixcr$meta, clonalityDF, by = "Sample")

##Plots
pdf(file="CD8_immdata_mixcr_trb_clonality.pdf")

vis(div_div) + scale_x_discrete(labels=CD8_immdata_mixcr$meta$Publication_id)+ggtitle("Clonality")+ylab("Clonality")+ylim(0,0.65)

vis(div_div, .by = "OS", .meta = CD8_immdata_mixcr$meta, .test = T) + ggtitle("Clonality by OS")+ylab("Clonality")+ylim(0,0.65)
#on plot p=0.09
wilcox.test(CD8_immdata_mixcr$meta$clonality ~ CD8_immdata_mixcr$meta$OS, exact = F)
#p=0.08273
t.test(CD8_immdata_mixcr$meta$clonality ~ CD8_immdata_mixcr$meta$OS, exact = F)
#0.02681

vis(div_div, .by = "OS", .meta = CD8_immdata_mixcr$meta, .test = F) + ggtitle("CD8 T cell Clonality by OS")+ylab("Clonality")+ylim(0,0.65) + stat_compare_means(comparisons = list(c("<2y",">2y")), method = "t.test")


vis(div_div, .by = "Treatment", .meta = CD8_immdata_mixcr$meta, .test = T) + ggtitle("Clonality by Treatment")+ylab("Clonality")+ylim(0,0.65)
#p=0.4
wilcox.test(CD8_immdata_mixcr$meta$clonality ~ CD8_immdata_mixcr$meta$Treatment, exact = F)
#p=0.4208
t.test(CD8_immdata_mixcr$meta$clonality ~ CD8_immdata_mixcr$meta$Treatment, exact = F)
#0.9548

vis(div_div, .by = "Treatment", .meta = CD8_immdata_mixcr$meta, .test = F) + ggtitle("CD8 T cell Clonality by Treatment")+ylab("Clonality")+ylim(0,0.65) + stat_compare_means(comparisons = list(c("GVAX_CY","GVAX_CY_NIVO")),method = "t.test")

vis(div_div, .by = "CD66b_Gr_Status", .meta = CD8_immdata_mixcr$meta) + ggtitle("Clonality by CD66b_Gr_Status")+ylab("Clonality")+ylim(0,0.65)
#p=0.8
wilcox.test(CD8_immdata_mixcr$meta$clonality ~ CD8_immdata_mixcr$meta$CD66b_Gr_Status, exact = F)
#p=0.7768
t.test(CD8_immdata_mixcr$meta$clonality ~ CD8_immdata_mixcr$meta$CD66b_Gr_Status, exact = F)
#0.6104
vis(div_div, .by = "CD66b_Gr_Status", .meta = CD8_immdata_mixcr$meta, .test=FALSE) + ggtitle("CD8 T cell Clonality by CD66b_Gr_Status")+ylab("Clonality")+ylim(0,0.65) + stat_compare_means(comparisons = list(c("High","Low")),method = "t.test")

vis(div_div, .by = "CD8_CD137_Status", .meta = CD8_immdata_mixcr$meta) + ggtitle("Clonality by CD8_CD137_Status")+ylab("Clonality")+ylim(0,0.65)
#p=0.06
wilcox.test(CD8_immdata_mixcr$meta$clonality ~ CD8_immdata_mixcr$meta$CD8_CD137_Status, exact = F)
#p=0.05183
t.test(CD8_immdata_mixcr$meta$clonality ~ CD8_immdata_mixcr$meta$CD8_CD137_Status, exact = F)
#0.008868
vis(div_div, .by = "CD8_CD137_Status", .meta = CD8_immdata_mixcr$meta, .test=F) + ggtitle("CD8 T cell Clonality by CD8_CD137_Status")+ylab("Clonality")+ylim(0,0.65) + stat_compare_means(comparisons = list(c("High","Low")),method = "t.test")

vis(div_div, .by = "CD4_Th17_Status", .meta = CD8_immdata_mixcr$meta) + ggtitle("Clonality by CD4_Th17_Status")+ylab("Clonality")+ylim(0,0.65)
#p=0.5
wilcox.test(CD8_immdata_mixcr$meta$clonality ~ CD8_immdata_mixcr$meta$CD4_Th17_Status, exact = F)
#p=0.5228
t.test(CD8_immdata_mixcr$meta$clonality ~ CD8_immdata_mixcr$meta$CD4_Th17_Status, exact = F)
#0.3469
vis(div_div, .by = "CD4_Th17_Status", .meta = CD8_immdata_mixcr$meta, .test = F) + ggtitle("CD8 T cell Clonality by CD4_Th17_Status")+ylab("Clonality")+ylim(0,0.65) + stat_compare_means(comparisons = list(c("High","Low")),method = "t.test")

vis(div_div, .by = "Treg_Status", .meta = CD8_immdata_mixcr$meta, .test = F) + ggtitle("CD8 T cell Clonality by Treg_Status")+ylab("Clonality")+ylim(0,0.65) + stat_compare_means(comparisons = list(c("High","Low")),method = "t.test")

vis(div_div, .by = "IL17_Signaling_Status", .meta = CD8_immdata_mixcr$meta, .test = F) + ggtitle("CD8 T cell Clonality by IL17_Signaling_Status")+ylab("Clonality")+ylim(0,0.65) + stat_compare_means(comparisons = list(c("High","Low")),method = "t.test")

dev.off()
```

## Extraction for DESEq comparison -- Continued in RNA scripts
```{r, include=FALSE}
##Extract CD8 #uniqueClones (Volume) and #clones/repertoire (clones)
cd8volume <- repExplore(CD8_immdata_mixcr$data, .method="volume")
cd8clones <- repExplore(CD8_immdata_mixcr$data, .method="clones")

##Extract CD4 #uniqueClones (Volume) and #clones/repertoire (clones)
cd4volume <- repExplore(CD4_immdata_mixcr$data, .method="volume")
cd4clones <- repExplore(CD4_immdata_mixcr$data, .method="clones")

#Add SampleNames to each
##CD8
###Extract SampleNames from VOLUME$Sample
filenames <- cd8volume$Sample
###Remove batch1 additions
samplenames = gsub("Plus.+","",filenames)
###Remove batch2 additions
samplenames = gsub("_.+","",samplenames)
samplenames = gsub("B.","",samplenames)
###SaveFinalSampleNames
samplenamesCD4 <- samplenames
###Add cd4Volume and clones w/ sampleNames
cd8TCR<- cbind(samplenames,cd8volume$Volume,cd8clones$Clones)
colnames(cd8TCR) <- c("SampleNames","Volume","Clones")

##CD4
###Extract SampleNames from VOLUME$Sample
filenames <- cd4volume$Sample
###Remove batch1 additions
samplenames = gsub("Plus.+","",filenames)
###Remove batch2 additions
samplenames = gsub("_.+","",samplenames)
samplenames = gsub("B.","",samplenames)
###SaveFinalSampleNames
samplenamesCD4 <- samplenames
###Add cd4Volume and clones w/ sampleNames
cd4TCR<- cbind(samplenames,cd4volume$Volume,cd4clones$Clones)
colnames(cd4TCR) <- c("SampleNames","Volume","Clones")

##Save for use in RNA scripts
save(cd8TCR,cd4TCR, file="tcrcloneinfo.rda")
```






